% Modern Single-Column Resume Template
% Design Philosophy: Clean, ATS-friendly layout with professional navy accents
% Font: Lato for modern, readable appearance
% Spacing: Generous white space for better readability and visual hierarchy

\documentclass[11pt,a4paper]{article}

% ============= PACKAGES =============
% inputenc is unnecessary on UTF-8 engines (XeLaTeX/LuaLaTeX)
% \usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lato} % Modern, professional sans-serif font
\usepackage[hidelinks]{hyperref} % Clean hyperlinks without colored boxes
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{tabularx}
\usepackage{multicol}
\usepackage[final]{microtype} % Better kerning & fewer broken ligatures

% Encoding & hyphenation safety for ATS/PDF export
\hyphenpenalty=10000
\exhyphenpenalty=10000
\tolerance=1000
\emergencystretch=3em

% ============= PAGE LAYOUT =============
% Optimized margins for single-column layout - balanced white space
\geometry{
  top=0.3in,
  bottom=0.5in,
  left=0.65in,
  right=0.65in,
  nofoot
}

% ============= COLOR SCHEME =============
% Professional navy palette - subtle yet distinctive
\definecolor{navyblue}{RGB}{25, 55, 100}      % Primary: Section headers
\definecolor{darkgray}{RGB}{60, 60, 60}       % Body text
\definecolor{lightgray}{RGB}{120, 120, 120}   % Secondary text

% ============= SECTION FORMATTING =============
% Navy-colored section headers with subtle underline for visual separation
\titleformat{\section}
  {\color{navyblue}\Large\bfseries\raggedright}
  {}{0em}
  {}
  [\color{navyblue}\titlerule]

\titlespacing*{\section}{0pt}{10pt}{4pt}

% ============= LIST FORMATTING =============
% Compact bullets with optimal spacing for content density
\setlist[itemize]{
  leftmargin=*,
  itemsep=1pt,
  topsep=3pt,
  parsep=0pt,
  partopsep=0pt,
  label=\textbullet
}

% ============= CUSTOM COMMANDS =============
% Consistent formatting for job titles and dates with proper alignment
\newcommand{\resumeheading}[4]{
  \noindent
  \begin{tabularx}{\textwidth}{@{}X r@{}}
    \textbf{\color{navyblue}#1} & \textit{\color{lightgray}#2} \\
    \textit{#3} & \textit{\color{lightgray}#4}
  \end{tabularx}\vspace{1pt}
}

% Project headings with tech stack emphasis
\newcommand{\projectheading}[2]{
  \textbf{\color{navyblue}#1} \hfill \textit{\color{lightgray}#2}
  \vspace{-2pt}
}

% Skill category formatting for clean organization with consistent wrapping
\newcommand{\skillcategory}[2]{
  \textbf{\color{navyblue}#1:} \raggedright #2
}

% ============= DOCUMENT START =============
\begin{document}
\pagestyle{empty} % Clean layout without page numbers

% ============= HEADER SECTION =============
% Centered name with contact info - optimized for ATS parsing
\begin{center}
  {\Huge\bfseries\color{navyblue}RAMITH K S} \\[4pt]
  {\Large\color{darkgray}Senior Backend Engineer} \\[8pt]
  \color{darkgray}
  \href{tel:+919481968808}{+91 9481968808} $|$ 
  \href{mailto:ramithgowdakundoor123@gmail.com}{ramithgowdakundoor123@gmail.com} $|$ 
  \href{https://www.linkedin.com/in/ramith-k-s}{LinkedIn: ramith-k-s} $|$ 
  \href{https://github.com/ramithks}{GitHub: ramithks}
\end{center}

% ============= PROFESSIONAL SUMMARY =============
\section{Professional Summary}
\color{darkgray}
{\small Senior Backend Engineer (4+ years) specializing in Python/FastAPI, Node.js, and distributed systems on AWS. Delivers scalable, low-latency microservices (sub-300ms P95), 4× query speedups, and platforms handling 10K+ daily requests. Strong in caching strategies, event-driven architecture, schema optimization, and leading teams to ship reliable production systems.}

% ============= PROFESSIONAL EXPERIENCE =============
\section{Professional Experience}
\resumeheading{Zolnoi Innovations}{October 2024 -- Present}{Senior Backend Engineer}{Bengaluru}
\vspace{2pt}
\textit{\small Tech Stack: FastAPI, PostgreSQL, Redis, AWS, Docker, APScheduler}
\vspace{2pt}
\begin{itemize}
\item Architected a multi-channel notification engine processing 5,000+ alerts/day with 99.5\% delivery reliability, reducing operational load and alert fatigue.
\item Consolidated 200+ alert streams into digest summaries, cutting user noise by 75\% and improving engagement.
\item Reduced API P95 latency from 800ms to 300ms (60\%) through Redis caching, optimized DB calls, and critical-path refactors.
\item Shipped 15+ production features across microservices; raised test coverage to 85\%, preventing 30+ regressions.
\item Achieved 4× faster queries via schema redesign, index tuning, and execution plan optimization.
\end{itemize}

\vspace{3pt}

\resumeheading{JustPoll (Streetlab Internet Pvt Ltd)}{October 2021 -- March 2025}{Founder \& CTO}{Bengaluru}
\vspace{1pt}å
\textit{\small Tech Stack: Node.js, TypeScript, PostgreSQL, Redis, AWS (ECS, S3), WebSockets, Docker}
\vspace{2pt}
\begin{itemize}
\item Architected and scaled TypeScript backend serving 500+ beta users; improved query performance by 40\% via schema optimization and caching.
\item Designed WebSocket infrastructure supporting real-time features with distributed state and connection lifecycle management.
\item Cut deployment time by 60\% (45 \textrightarrow{} 15 mins) by introducing ECS autoscaling, CI/CD, and CloudWatch-driven observability.
\item Built Redis-backed task queues with <200ms P95 latency, reducing incidents by 45\%.
\item Led a 5-member engineering team, driving 20+ releases and improving user retention by 65\%.
\end{itemize}

\vspace{2pt}

% ============= TECHNICAL SKILLS =============
\section{Technical Skills}

\skillcategory{Backend \& Languages}{Python, FastAPI, Node.js, TypeScript, AsyncIO, REST, gRPC, WebSockets}

\vspace{2pt}

\skillcategory{Databases}{PostgreSQL, Redis, MongoDB, MySQL, Query Optimization, Indexing, Schema Design, ACID Transactions}

\vspace{2pt}

\skillcategory{Systems \& Architecture}{Microservices, Distributed Systems, Event-Driven Architecture, Caching Strategies, Load Balancing, High Availability, Rate Limiting, API Gateway, Observability}

\vspace{2pt}

\skillcategory{Cloud \& DevOps}{AWS (ECS, EC2, S3, CloudWatch, Lambda), Docker, Terraform, GitHub Actions, CI/CD, Infrastructure Automation}

\vspace{2pt}

\skillcategory{Testing \& Reliability}{Pytest, Integration Testing, SLO/SLI Monitoring, Incident Management, Root Cause Analysis}

\vspace{2pt}

% ============= EDUCATION =============
\section{Education}

\textbf{\color{navyblue}Bachelor of Computer Applications} \hfill \textit{\color{lightgray}2018 -- 2021}\\
Christ University, Bangalore \hfill CGPA: 3.3/4.0

\vspace{2pt}

% ============= RECOGNITION =============
\section{Recognition}

\begin{itemize}
\item HackBangalore Winner 2024
\item AWS Cloud Foundations Certified
\item Open Source Contributor
\end{itemize}

\vspace{2pt}

% ============= KEY PROJECTS =============
\section{Key Projects}

\projectheading{AlgoHabit | AI-Powered Learning Platform}{\href{https://algohabit.com}{algohabit.com} | 200+ Active Users}
\vspace{1pt}
\textit{\small Tech Stack: React, TypeScript, Supabase, OpenRouter AI, Razorpay, PWA}
\vspace{2pt}
\begin{itemize}
\item Engineered React PWA achieving 95+ Lighthouse performance score with offline-first architecture using service workers for 100\% offline functionality.
\item Integrated OpenRouter AI with intelligent caching layer reducing API costs by 70\%, processing 100+ user-uploaded syllabi with 90\% accuracy.
\item Deployed Supabase backend with RLS securing 5,000+ user records; enabled zero-downtime Razorpay payments.
\item Implemented gamification tracking 10,000+ XP transactions, improving 30-day retention by 60\%.
\end{itemize}

\vspace{7pt}

\projectheading{Ledgerly | Smart Finance Tracker SaaS}{Personal Project}
\vspace{1pt}
\textit{\small Tech Stack: FastAPI, React, PostgreSQL, ARIMA, TF-IDF, Pytest}
\vspace{2pt}
\begin{itemize}
\item Built FastAPI backend with ARIMA time-series forecasting predicting expenses 3 months ahead with 82\% accuracy.
\item Implemented TF-IDF-based transaction categorization achieving 88\% classification accuracy across 50+ categories.
\item Developed pytest suite covering 40+ endpoints with 92\% code coverage.
\end{itemize}

\vspace{7pt}

\projectheading{AG-WALLET | Agricultural Supply Chain Platform}{Android Application}
\vspace{1pt}
\textit{\small Tech Stack: Android (Java), MySQL, Google Maps API}
\vspace{2pt}
\begin{itemize}
\item Built Android app connecting farmers, distributors, and customers for transparent pricing and direct market access.
\item Implemented real-time price monitoring enabling farmers to track distributor pricing and maximize profit margins.
\item Integrated Google Maps API for location-based distributor discovery and rating system for quality-based selection.
\end{itemize}

\vspace{7pt}

\projectheading{HelpMyBuddy | HackBangalore Winner 2024}{\href{https://github.com/SiliconNinjas/HelpMyBuddy}{GitHub Link}}
\vspace{2pt}
\begin{itemize}
\item Developed React Native prototype connecting homeowners with service providers via geolocation matching within 5km.
\item Built Node.js backend with booking system and real-time availability tracking.
\end{itemize}

\end{document}